\documentclass{article}
\begin{document}
\centerline{\bf \Large Application for financial support for}
\centerline{\bf \Large Formal power series and algebraic combinatorics}
\centerline{July 20-24, 2009, Hagenberg, Austria}
\medskip
\noindent Deadline for application: March 1, 2009
\medskip\\
\noindent The conference expects to provide limited financial support
to contribute to travel and living expenses during the
conference. First priority will be given to junior faculty, graduate
and postdoctoral students, and, in some exceptional cases, to
qualified undergraduates. {\bf You must be affiliated with a US
institution to use this form.} Please send it to Susanna Fishel at
{\tt fpsacfunding@yahoo.com}. Please send the .pdf or .ps of
this. Please arrange to have your recommendation sent here too.
\medskip\\
\noindent If you are interested in support, please complete and submit
the following form.
\medskip
\begin{description}
\item[Name:]  

\item[US residency status:]

\item[E-mail Address:]

\item[Mailing Address:]

\item[University/Organization:]

\item[Current Position:]

\item[Year of Ph.D.:]

\item[University at which you did your Ph.D.:]
\end{description}
\begin{enumerate}
\item If you are a student, please indicate the name and email address of
your faculty supervisor.

\item Recommender's name and contact information:

\item Estimated transportation cost and living expenses:
 
\item Please list all other sources of financial support:

\item Are you presenting a talk or poster?

\item Please explain in a few sentences why you need the support and how you would expect to benefit from the conference.
\end{enumerate}
\end{document}
